---
title: "3. Plots"
author: "Chris Terry"
date: "28/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE,  warning = FALSE}
library(tidyverse)
library(lme4)
library(cowplot)
library(ggraph)
library(tidygraph)
library(igraph)
```

# Loading and unpacking results outputs 

Saved as a list of dataframes

```{r}
load( 'Outputs/Lags' )
load( 'Outputs/CritCC')

ParamDF_NC <- read_csv('Outputs/ParamDF_NC.csv')

Lags%>%
  map('result')%>%
  compact %>%
  map_df(~.) %>%
  left_join(ParamDF_NC, by = "RunID")->Lags_Results

CritCC%>%
  map('result')%>%
  compact %>%
  map_df(~.) %>%
  left_join(ParamDF_NC, by = "RunID")-> CritCC_DF

```

# Critical CC Values

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
### NB Joining the two df of Critical CC Values together

CritCC_DF %>%
 mutate(FracExtant1 = NumberNowExtant_1  /NumberFocalStart ,
        FracExtant001 = NumberNowExtant_0.01  /NumberFocalStart) -> AllCritCC_df

AllCritCC_df%>%
  ggplot( aes( col = factor(EPC_MorseA)  , 
               y = FracExtant1, x =CC_rate  ))+
  geom_point()+
  scale_color_viridis_d()+
  facet_grid(TempNoise~mRate, labeller = label_both)+
  geom_smooth(se=FALSE, method = 'glm' , 
              method.args=list(family="binomial"))+
  ggtitle( 'Across all movement rates')

```

## CC critical speed stats

```{r}
JustSlowData <- filter( AllCritCC_df, mRate == 1e-07)

Full_RE <- glmer( data = JustSlowData, family = 'binomial',  weights = NumberFocalStart  ,
                  FracExtant1 ~   CC_rate*factor(EPC_MorseA)  + (1|RunID  ) ) 

summary(Full_RE)

## shifts 50% extinction rate point by :
Coefficents<-summary(Full_RE)$coef[,1]

#Prediction for unscaled y = 0 (frac extinct prob =0.5)

## With asym = -1:

## 0 = intercept - CC_rate C*
##   C* = - intercept / CC_Rate
(-Coefficents[1] )/ Coefficents[2]

## With asym = 1:
## 0 = intercept + CC_rate C*   + Assym + CC_rate x interaction x C*
##   C* = - (intercept+ Assym) / (CC_Rate + interaction)  )
(-(Coefficents[1] + Coefficents[3] ))/ (Coefficents[2]+ Coefficents[4])

```

# Lags

```{r echo=FALSE}
Lags_Results %>%
  ggplot(aes( x = SpMeanLag , col = factor(EPC_MorseA)))+
  geom_density( )+
  scale_colour_viridis_d()+
  facet_grid(TempNoise  ~ mRate , 
             labeller = label_both)+
  ggtitle( 'Effect of Assymetry direction')


Single_Lags_Full <- lmer( data = Lags_Results,
                           SpMeanLag  ~ factor(EPC_MorseA)* log10(mRate) + (1|RunID))

summary(Single_Lags_Full)


Single_Lags <- lmer( data = filter( Lags_Results, mRate == 1e-07),
                     SpMeanLag  ~ factor(EPC_MorseA)+ (1|RunID))

summary(Single_Lags)


filter( Lags_Results, mRate == 1e-07) %>%
group_by(EPC_MorseA) %>%
    summarise( mean(SpMeanLag, na.rm=TRUE ))

```

# Combined plot for paper

### demonstration plot

```{r}
## Load spoPool from previously generated assembly

load('AssemblyExample/sppPool_ID_1')

## Use function 

source( 'FunctionScripts/R functions_discete.R')


EdgeList<-as_tbl_graph(graph_from_adjacency_matrix(sppPool$topo$adjMat,
                                                   diag = FALSE, mode = 'undirected'))

EdgeList %>%
  activate(nodes) %>%
  mutate( Node_X = sppPool$network[,1], 
          Node_Y = sppPool$network[,2],
          Env = as.vector(sppPool$envMat),
          SpeciesDensity = log10(sppPool$bMat_p[1,])) -> GraphObject

Curve <- data.frame( Env = seq(10, 40, l = 1000),
            Response = CalcR_Morse( seq(10, 40, l = 1000),
                                    opt = sppPool$tVec[1], 1, -1, 10, -100 )) %>%
  filter( Response>0.1) %>%
  mutate( ScaleResponse = (Response /3)+11 )

GraphObject %>%
  ggraph(layout = data.frame(x=sppPool$network[,1],
                             y=sppPool$network[,2])) + 
  geom_edge_link() + 
  geom_node_point(aes(fill = Env), size = 4, shape = 21)+
  geom_node_point(aes(size = SpeciesDensity), shape = 23,
                  fill = 'yellow', alpha = 0.5)+
  coord_fixed()+
  scale_fill_gradient2(low = 'blue1',midpoint = 20, 
                       mid = 'beige', high = 'tomato3')+
  #scale_fill_viridis_c( ) +
  theme(rect = element_blank())+
  geom_line( data = Curve, aes( x = Env, y = ScaleResponse), size = 2)+
  guides( size = FALSE, fill = FALSE)+
  annotate("segment", x = 20, xend = 15, y = 12, yend = 12,
           colour = "black", size=3, arrow=arrow())-> ArenaPlot

ArenaPlot


GraphObject %>%
  ggraph(layout = data.frame(x=sppPool$network[,1],
                             y=sppPool$network[,2])) + 
  geom_edge_link() + 
  geom_node_point(aes(fill = Env), size = 4, shape = 21)+
  coord_fixed()+
  scale_fill_gradient2(low = 'blue1',midpoint = 20, 
                       mid = 'beige', high = 'tomato3')+
  theme(rect = element_blank())+
  guides( size = FALSE, fill = FALSE)-> SimpleArenaPlot

SimpleArenaPlot
ggsave( 'Figures/SimpleArenaPlot.png', SimpleArenaPlot,width = 8, height = 3, dpi = 500)

```

```{r}

AllCritCC_df%>%
  filter( mRate == 1e-07 ) %>%
  ggplot( aes( col = factor(EPC_MorseA)  , 
               y = FracExtant1, x =CC_rate  ))+
  geom_point()+
  geom_smooth(se=FALSE, method = 'glm' , 
              method.args=list(family="binomial"))+
  ylab('Fraction of Species Extant')+
  xlab('Rate of climate change (v)')+
  geom_hline(yintercept = 0.5)+
  theme_classic()+
  ggtitle( 'i) Critical rate of climate change')+
  scale_colour_viridis_d(name = 'Asymmetry parameter:',  )+
  theme(axis.text.y = element_blank(), 
        legend.direction = 'horizontal')-> PlotA


Lags_Results %>%
    filter( mRate == 1e-07 ) %>%
  ggplot(aes( x = SpMeanLag , col = factor(EPC_MorseA)))+
  geom_density( )+
  ggtitle( 'ii) Lag behind climate shifts')+
  xlab('Species Lag')+
    scale_colour_viridis_d(name = 'Asymmetry\nparameter',  )+
  ylab('')+
  theme_classic()+
  guides(col =FALSE)-> PlotB



results_noleg<-plot_grid(PlotA+guides(col =FALSE), PlotB,
                         rel_widths = c( 2,2), nrow = 1, scale = 0.9)

ResultsPlots<- plot_grid(results_noleg,get_legend(PlotA),
                         ncol = 1, rel_heights = c(1,0.05))


```

# Illustration of EPCs

```{r}

 data.frame( Env = seq(-2, 7, l = 1000),
                      Right = CalcR_Morse( seq(-2, 7, l = 1000),
                                           opt = 0, 1, -1, 10, -100 ),
                      Left = CalcR_Morse( seq(-2, 7, l = 1000),
                                          opt = 5, 1, 1, 10, -100 )) %>%
  gather( key = 'Assymetry', value = 'R', - Env) %>%
  filter( R > -1) %>%
  mutate( Assymetry =  ifelse(Assymetry=='Right',
                              'Right-skewed\n(a = -1)',
                              'Left-skewed\n(a= +1)'  )) %>% 
  ggplot() +
  geom_line( aes( x = Env, y = R, col = Assymetry), size = 3)+
  facet_grid(Assymetry~.)+
  theme_classic()+
  scale_colour_viridis_d(direction = -1)+
  guides( colour = FALSE)+
  theme(strip.background = element_blank(),
        axis.text = element_blank(),
        axis.line = element_blank(),
        axis.title  = element_blank(),
        axis.ticks  = element_blank()) -> PlotC

 PlotC
 
 ### Making colour bar
colourbar_plot <- ggplot(data = data.frame(x = 0:10,
                                           y = 0:10),
                         aes(x = x, y = y,
                             colour = y)) + 
  geom_point() +
  scale_color_gradient2(low = 'blue1',midpoint = 5,
                       mid = 'beige',
                       high = 'tomato3',
                       guide =guide_colourbar(direction = "horizontal",
                                                 barwidth = 10,
                                                  title = element_blank(),
                                                 label = FALSE))+
  theme_minimal()

 PlotC_bar <- plot_grid(PlotC, get_legend(colourbar_plot),
                        ncol = 1, rel_heights = c( 3,1))
 
 
### Combining
# plot_grid(PlotA, PlotB,PlotC_bar, rel_widths = c( 2,2,1), nrow = 1, scale = 0.85) -> ResultsPlots


```

# All together to make Results Figure


```{r}
SimulationResults<- plot_grid(ArenaPlot, ResultsPlots, ncol = 1,
                              labels = c('a) Diagram of Arena', 
                                         'b) Results'),
                              hjust = -0.1, label_fontface = 'plain' )
SimulationResults
ggsave('Figures/SimulationResults.png', SimulationResults,
       dpi = 600, height = 6, width =10)
```

# EPC Figure

## Harmonic potential

$$ r =  R_{max}*\left(1- \frac{(E_{vec}-opt)^2}{w^2} \right)$$ 

```{r}
CalcR_Harm<- function( E_vec, opt, EPC_w,  R_Max, EPC_Min ){
  y = E_vec-opt  
  R <-  (R_Max * (1 -   ((y^2)/(EPC_w^2))  )  ) 
  R[R<EPC_Min] <- EPC_Min
  return(R)
}

```

### Basic Morse potential:

simplifying distance from the optima $ y = E_{vec}-opt$

$$R_{Max} * (1 - exp(- a y))^2)$$

### Morse Potential with controllable width

This also include a a niche width term $w$ , and 
$$  R_{Max}(1-\frac{(1-e^{ay})^2}{a^2w^2})$$
where:

- $r_{max}$ is the peak
- $a$ is the assymety
- $w$ is the niche width
- $ y = E_{vec}-opt$
- $E_{opt}$ = the location of the peak


```{r}
CalcR_Morse<- function( E_vec, opt, EPC_w, EPC_MorseA, R_Max, EPC_Min ){
  y = E_vec-opt
  R <-  R_Max * (1- ((1 - exp(EPC_MorseA *y))^2)/ (EPC_MorseA^2 * EPC_w^2) ) 
  R[R<EPC_Min] <- EPC_Min
  return(R)
}

```

# Figure of EPCs for paper

```{r}

R_Max = 10
E_vec = seq(-10, 10, by = 0.001)

A_Harmonic <- data.frame(E_vec = E_vec,
                         R = CalcR_Harm(E_vec, opt = 0, EPC_w =2 , R_Max =R_Max, EPC_Min= -10 ),
                         EPC = 'Harmonic')

B_MorsePos <- data.frame(E_vec = E_vec,
                         R=  CalcR_Morse(E_vec,  opt = 0, EPC_w = 1, 
                                         EPC_MorseA = 1, R_Max =10,
                                         EPC_Min = -10 ),
                         EPC = 'Morse_Pos' )

C_MorseNeg <- data.frame(E_vec = E_vec,
                         R= CalcR_Morse(E_vec,  opt = 0, EPC_w = 1, 
                                          EPC_MorseA = -1, R_Max =10,
                                          EPC_Min = -10 ),
                         EPC = 'Morse_Neg' )

#### Convolving with weather:
f.X <- function(x) dnorm(x,0,1)  ## weather of sd = 1 
f.G <- function(G) CalcR_Morse(G,  opt = 0, EPC_w = 1, EPC_MorseA = 1, R_Max =10, EPC_Min = -10 ) # basis
f.Z2 <- function(z) integrate(function(x,z) f.G(z-x)*f.X(x),-Inf,Inf,z)$value
f.Z2 <- Vectorize(f.Z2)                    # need to vectorize the resulting fn.
 

D_MorsePosWeather <- data.frame(E_vec = E_vec,
                                R = f.Z2(E_vec),
                                EPC = 'Morse_Pos_Weather')

```

```{r}
          
A_Harmonic_Plot <-           A_Harmonic %>%
  ggplot( aes( E_vec, R))+
  geom_line(size = 1)+
  coord_fixed(ylim = c(0.1, 10), xlim = c( -5,5))+
  theme_classic()+
  geom_vline(xintercept = 0, linetype = 'dashed')+ xlab('E')


B_MorsePos_Plot <-   B_MorsePos %>%
  ggplot( aes( E_vec, R))+
  geom_line(size = 1)+
  coord_fixed(ylim = c(0.1, 10), xlim = c( -5,5))+
  theme_classic()+
  geom_vline(xintercept = 0, linetype = 'dashed')+ xlab('E')

C_MorseNeg_Plot <-C_MorseNeg %>%
  ggplot( aes( E_vec, R))+
  geom_line(size = 1)+
  coord_fixed(ylim = c(0.1, 10), xlim = c( -5,5))+
  theme_classic()+
  geom_vline(xintercept =0, linetype = 'dashed')+ xlab('E')

D_MorsePosWeather_Plot <-D_MorsePosWeather %>%
  ggplot( aes( E_vec, R))+
  geom_line(size = 1)+
  coord_fixed(ylim = c(0.1, 10), xlim = c( -5,5))+
  theme_classic()+
  geom_vline(xintercept = 0, linetype = 'dashed')+ xlab('E')


plot_grid(A_Harmonic_Plot, B_MorsePos_Plot,
          C_MorseNeg_Plot,D_MorsePosWeather_Plot,
          nrow = 2, align = 'hv',
          labels = c('a) Harmonic',
                     'b) Morse (a = +1)',
                     'c) Morse (a = -1)',
                     'd) With underlying variation' ) ,
          hjust = -0.1, label_fontface = 'plain', scale = 0.85) -> EPCexamples

EPCexamples

ggsave('Figures/EPCexamples.png', width = 6, height = 5, dpi = 600)

```

