---
title: "2 Simulation Run"
author: "Chris Terry"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document runs the code to assemble the communities, then run two tests of the impact of climate change 

It's main output is two lists of data.frames, containing the results of the two tests. These are included in the repository. 

The assembled communities are large (1.2Mb each) so are not included other than an example. If wanted, they can be generated from the code here. 

```{r message = FALSE}
## Load packages
library(tidyverse)

### Load all functions
walk( list.files('FunctionScripts/', full.names = TRUE), source)

```

# Assembling communities
## Control Parameter setup

Testing two different dispersal rates, but only show one in the results.

```{r}
TrialName  = 'Morse'

N_Repeats = 100  # repeats per run parameter combination. 
ParamDF_NC <- expand.grid(  Repeat= paste0('R_', 1:N_Repeats),
                            mRate = c(1e-7, 1e-5),
                            EPC_MorseA = c( -1, 1),
                            TempNoise=  c(0.5),
                            stringsAsFactors = FALSE) %>%
  mutate(RunID =  paste0('ID_' , row_number()),
         TrialName = TrialName)

write_csv(ParamDF_NC, file = 'Outputs/ParamDF_NC.csv')  # saving 'run' parameters

```

## Running Assembly

Map over each assembly, outputting assembled communities. Takes about 20 seconds per community on an average desktop. 1 example is included. 


```{r eval = FALSE}
NC_TrialAssm <- map(1:nrow(ParamDF_NC),
                    PreAssembly, ParamList = ParamDF_NC,
                    num_species =100,
                    length_Relax= 200,
                    ParameterSource = 'Parameters/Parameters.R')
```
# Testing Lags


```{r eval = FALSE}

Lags<-map(1:nrow(ParamDF_NC),
                   safely(Lag_Measurer),
                   ParamList = ParamDF_NC,  
                   CC_rate= 0.1, # To match analytics, using quite a slow rate of climate change
                   length_preCC = 20,
                   length_CC_preTest = 50,
                   length_CC=20)

 # save(Lags, file = 'Outputs/Lags' )

```

# Testing Crit CC

```{r eval = FALSE}
CritCC<-map(1:nrow(ParamDF_NC),
                        safely(CritCC_Calculator),
                        ParamList = ParamDF_NC,  
                        length_CC=50,
                        CC_rates= seq(0,0.5, l = 11)) ## rates of climate change to use

# save( CritCC, file = 'Outputs/CritCC')
```

# System Information 


```{r}
sessionInfo()
```



